### 2018-07-11 submission/manuscript.Rmd
add fecal to biological variation

Previous studies have reported GC dependent amplification differences [@polz_bias_1998] and we confirm that these differences still persist in high fidelity DNA polymerase when analyzing mock community samples [Figure 1]. Additionally, these specific polymerases have different sequence error rates and chimera prevalence [Figure 2]. Chimera prevalence in particular was strongly associated with the total number of OTUs detected in a mock community [Figure 2E]. This total number of OTUs also varied by polymerase used for these mock community samples [Fiugre 3A]. The range in the number of OTUs observed increased with the number of cycles and the change in range was dependent on the polymerase for both mock community and fecal samples [Figure 3]. However, these differences did not translate to large bacterial community composition changes based on polymerase [Figure 4]. Even though there were bacterial community composition changes based on the number of cycles they occured independently of the polymerase used [Figure 4]. Additionally, there were no OTU-based differences observed within the number of cycles between any of the polymerases for fecal samples. When Random Forest models were created to classify number of cycles, polymerase, and individual based on the 16S rRNA gene sequencing, the model performance for individuals was significantly greater than all other models tested [Figure 5A]. In addition to the better model performance of individuals, the most important OTUs to this model were not the most important OTUs for the other models tested [Figure 5B]. Our observations show that differences between high fidelity DNA polymerases can be found. However, when using 16S rRNA gene sequencing these technical differences are dwarfed by biological differences between individuals.

@@

Collectively, our data shows that using distinct high fidelity DNA polymerases will result in a different number of OTUs being detected. The differences in the number of OTUs are primarily due to chimera prevalence and sequence error rates that are distinct to the specific polymerase. Despite this variation in the number of OTUs, our Random Forest models were able to correctly classify individuals regardless of the number of cycles and polymerase used. These results suggest that even though there is a large variation in the high fidelity DNA polymerases used across studies, biological variation similar in effect to bacterial community composition differences between individuals will still be consistent.

@@

Although care should always be taken when choosing a polymerase for 16S rRNA gene sequencing, our observations show that the differences between a variety of polymerases are dwarfed by the actual biological variation between individuals. Therefore, if the biological signal of interest is similar to differences in bacterial community composition found between individuals, then the type of high fidelity DNA polymerase used will only minimally change the results.

>>>>

Previous studies have reported GC dependent amplification differences [@polz_bias_1998] and we confirm that these differences still persist in high fidelity DNA polymerase when analyzing mock community samples [Figure 1]. Additionally, these specific polymerases have different sequence error rates and chimera prevalence [Figure 2]. Chimera prevalence in particular was strongly associated with the total number of OTUs detected in a mock community [Figure 2E]. This total number of OTUs also varied by polymerase used for these mock community samples [Fiugre 3A]. The range in the number of OTUs observed increased with the number of cycles and the change in range was dependent on the polymerase for both mock community and fecal samples [Figure 3]. However, these differences did not translate to large bacterial community composition changes based on polymerase [Figure 4]. Even though there were bacterial community composition changes based on the number of cycles they occured independently of the polymerase used [Figure 4]. Additionally, there were no OTU-based differences observed within the number of cycles between any of the polymerases for fecal samples. When Random Forest models were created to classify number of cycles, polymerase, and individual based on the 16S rRNA gene sequencing, the model performance for individuals was significantly greater than all other models tested [Figure 5A]. In addition to the better model performance of individuals, the most important OTUs to this model were not the most important OTUs for the other models tested [Figure 5B]. Our observations show that differences between high fidelity DNA polymerases can be found. However, when using 16S rRNA gene sequencing these technical differences are dwarfed by biological differences between fecal bacterial communities within individuals.

@@

Collectively, our data shows that using distinct high fidelity DNA polymerases will result in a different number of OTUs being detected. The differences in the number of OTUs are primarily due to chimera prevalence and sequence error rates that are distinct to the specific polymerase. Despite this variation in the number of OTUs, our Random Forest models were able to correctly classify individuals regardless of the number of cycles and polymerase used. These results suggest that even though there is a large variation in the high fidelity DNA polymerases used across studies, biological variation similar in effect to fecal bacterial community composition differences between individuals will still be consistent.

@@

Although care should always be taken when choosing a polymerase for 16S rRNA gene sequencing, our observations show that the differences between a variety of polymerases are dwarfed by the actual biological variation in fecal communities between individuals. Therefore, if the biological signal of interest is similar to differences in fecal bacterial communities found between individuals, then the type of high fidelity DNA polymerase used will only minimally change the results.
